---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setglobaloption,echo=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```

## Loading and preprocessing the data

Show any code that is needed to

1. Load the data (i.e. read.csv())
```{r load_data}
unzip("activity.zip")
data<-read.csv("activity.csv");
```

2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r process_data}
library(lubridate)
data$date<-ymd(data$date)
```


## What is mean total number of steps taken per day?

For this part of the assignment, you can ignore the missing values in the dataset.

1. Calculate the total number of steps taken per day:
```{r total_steps}
totalSteps<-tapply(data$steps,data$date,sum,na.rm=TRUE)
```

2. If you do not understand the difference between a histogram and a barplot, research the difference 
between them. Make a histogram of the total number of steps taken each day

```{r hist_totSteps}
hist(totalSteps)
```

3. Calculate and report the mean and median total number of steps taken per day

```{r mean_median_steps,results='asis'}
library(xtable)
meanSteps<-tapply(data$steps,data$date,mean,na.rm=TRUE)
medianSteps<-tapply(data$steps,data$date,median,na.rm=TRUE)
dfSteps<-data.frame(date=names(meanSteps),mean=meanSteps,median=medianSteps,row.names = NULL)
xt<-xtable(dfSteps);
print(xt,type="html")
```

## What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number 
 of steps taken, averaged across all days (y-axis)
```{r meanInterval}
intervals<-unique(data$interval)
meanIntervalSteps<-tapply(data$steps,data$interval,mean,na.rm=TRUE);
plot(intervals,meanIntervalSteps,type = "l")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number 
 of steps?

The interval which, on avarage, contains the maximum number of steps, i.e. `r max(meanIntervalSteps)` 
is the interval number `r which.max(meanIntervalSteps)[[1]]`, corresponding to interval 
`r names(meanIntervalSteps)[which.max(meanIntervalSteps)[[1]]]`



## Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA). 
The presence of missing days may introduce bias into some calculations or summaries of the data.

1. Calculate and report the total number of missing values in the dataset 
(i.e. the total number of rows with NAs)

```{r NA}
totalNA<-sum(is.na(data$steps))
```

The number of NA values in the data is `r totalNA `.

2. Devise a strategy for filling in all of the missing values in the dataset. 
 The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, 
 or the mean for that 5-minute interval, etc.
 
NA values are replaced with the mean of the corresponding interval over all days.

```{r NAreplacement}
NAreplacement<-data.frame(steps=data$steps,interval=data$interval)
for (i in seq(1,nrow(data))){
   if (is.na(NAreplacement$steps[i])){
      NAreplacement$steps[i]<-meanIntervalSteps[which(NAreplacement$interval[i]==unique(NAreplacement$interval[i]))]
   }
}
```

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.


```{r dataCorrected}
dataCorrected<-data;dataCorrected$steps<-NAreplacement$step
sum(is.na(dataCorrected$steps))
```

4. Make a histogram of the total number of steps taken each day and Calculate and report 
  the mean and median total number of steps taken per day. 
  

```{r valuesDataCorrectedhist}
totalStepsCorrected<-tapply(dataCorrected$steps,dataCorrected$date,sum,na.rm=TRUE);
hist(totalStepsCorrected)
```

```{r valuesDataCorrected,results="asis"}
library(xtable)
meanCorrected<-tapply(dataCorrected$steps,dataCorrected$date,mean);
medianCorrected<-tapply(dataCorrected$steps,dataCorrected$date,median);
dfCorrected<-data.frame(date=names(meanCorrected),mean=meanCorrected,median=medianCorrected,row.names = NULL)
print(xtable(dfCorrected),type="html")
```


Do these values differ from the estimates from the first part of the assignment? 
  What is the impact of imputing missing data on the estimates of the total daily number of steps?
 
We can see from the histograms below that the total number of steps for each day doesn't change when 
NA are replaced.

```{r comparison1} 
dates<-as.Date(attributes(meanSteps)$dimnames[[1]])

par(mfrow=c(2,1),mar=c(4,4,1,1))
hist(totalSteps)
hist(totalStepsCorrected)
```

On the other hand, there is a small increase for the mean and a significant increase on the median.

```{r comparison2}
par(mfrow=c(2,1),mar=c(4,4,1,1))
plot(dates,meanSteps,type="l",col="blue");
plot(dates,meanCorrected,type="l",col="red")

par(mfrow=c(2,1),mar=c(4,4,1,1))
plot(dates,medianSteps,type="l",col="blue");
plot(dates,medianCorrected,type="l",col="red")
```

## Are there differences in activity patterns between weekdays and weekends?

For this part the weekdays() function may be of some help here. Use the dataset with the 
filled-in missing values for this part.

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” 
 indicating whether a given date is a weekday or weekend day.

```{r week}
library(lubridate)
dataCorrected$date<-ymd(dataCorrected$date)
weekday<-weekdays(dataCorrected$date)
weekday<-factor(weekday,levels=c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"))
dataCorrected$week<-rep(0,length(weekday))
for (i in seq(1,length(weekday))){
   if (weekday[i] %in% c("Monday","Tuesday","Wednesday","Thursday","Friday")){
      dataCorrected$week[i]<-"weekday"
   }
   else {dataCorrected$week[i]<-"weekend"}
}

dataCorrected$week<-factor(dataCorrected$week)
```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval 
 (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
 See the README file in the GitHub repository to see an example of what this plot should look like 
 using simulated data.

```{r weekday_end}
dataCorrectedWeek<-split(dataCorrected,dataCorrected$week)
meanStepsWeekday<-tapply(dataCorrectedWeek[[1]]$steps,dataCorrectedWeek[[1]]$interval,mean);
meanStepsWeekend<-tapply(dataCorrectedWeek[[2]]$steps,dataCorrectedWeek[[2]]$interval,mean);

par(mfrow=c(2,1))
plot(intervals,meanStepsWeekday,type = "l",ylab="weekday",main="Averaged number of steps")
plot(intervals,meanStepsWeekend,type = "l",ylab="weekend")
```